\noindent\rule{13cm}{1.0pt}\\ \vspace{0.2cm} A Polynomial-time Reduction of the 3-SAT to the Quadratic Congruence and other Related Problems \noindent\rule{13cm}{0.8pt}

### 2017-11-21 paper/paper.tex
Reviewed version with Intuition.

From this we know that one can reduce 3-SAT to any problem that the latter will be proved to be
NP-complete. In the next sections we will provide a brief review on polynomial-time reductions,
present proper definitions on the QCP and 3-SAT, and finally prove the reduction. The last section
is devoted to related problems. We show the Diophantine problem presented in~\cite{qcp2} and other
Number Theory problems reductions.

@@

>>>>

From this we know that one can reduce 3-SAT to any problem that the latter will be proved to be
NP-complete. In the next sections we will provide a brief review on polynomial-time reductions,
present proper definitions on the QCP and 3-SAT, and finally prove the reduction. We then try to
give some intuition on the proof. The last section is devoted to related problems. We show the
Diophantine problem presented in~\cite{qcp2} and other Number Theory problems reductions.

@@

\section{Intuition}

In this section we try to give some intuition on the reduction. It might be convenient to keep the
reduction and algorithm sections close by, so that one can reference them back following the
statements made here.

One can summarize the proof as trying to find an algorithm that converts a 3-CNF formula to an
expression of the quadratic congruence form. In the algorithm stated in the last section, we took
as input the 3-CNF formula $\phi$ and as output an expression of the QCP form. We do not give a
particular form for $\phi$ because it is not needed for the proof. In fact, we only need to prove
that $\phi$ is satisfiable if and only if the quadratic congruence formula is satisfiable (i.e.\
there is a solution for $x$ obeying the QCP conditions). The satisfiability of $\phi$ is guaranteed
by the 3-SAT problem, so our goal is then to show that if we can find a valuation for the 3-SAT
problem, then there is a corresponding valuation for the QCP to be satisfiable.

The proof's first step is to show that, by using a certain function $r$, we can transform the 3-CNF
formula to a congruence. That is, show that the LHS below is satisfiable if and only if the RHS is
satisfiable.

\begin{equation*}
  \sum_{k=0}^m R_k\cdot 8^k\equiv 0\mod 8^{m+1} \iff \phi
\end{equation*}

We do this by defining $R_k$ dependent on $\sigma_k$, so that $R_k$ is satisfiable if and only if
all $\sigma_k$ are satisfiable (and therefore $\phi$ satisfiable). Now that the set of all $R_k$ is
satisfiable if and only if $\phi$ is satisfiable, we need to manipulate $R_k$ such that there is a
congruence in there somewhere, as we need to have something that resembles $x^2\alpha\equiv\beta$.
We do this by summing all clauses $\sigma_k$. Note that in~\autoref{eq-2}, we are summing all
$\sigma_k$ through $R_k$. In fact, one can think of each term $R_k\cdot 8^k$ as an octal encoding
of each 3-clause ($2^3$) (thanks to Thales for this great insight). This encoding maps a certain
valuation $y_k$ at each clause. If we look back to the proof of~\autoref{eq-1}, we can see that
there are 4 possible valuations (one can easily see that other permutations are equivalent) for
$y_k$. This valuation is encoded by $y_k$, as there are only 4 possible values of $y_k$ that make
$R_k$ satisfiable: $\{0,1,2,3\}$. So the set of $\{y_k\in\{0,1,2,3\}|0\leq k\leq m\}$ is actually
the 3-SAT valuation.  Note how in~\autoref{eq-2}, for $R_k$ to be zero, $y_k\in\{0,1,2,3\}$, and so
the summation on the RHS encodes the same valuation.

The goal is then to keep this encoding satisfiable if and only if $\phi$ is satisfiable, and also
get something of the form $x^2\equiv\alpha\mod\beta$. We do this by getting to~\autoref{eq-3}.
\autoref{eq-p2} then replaces variables from~\autoref{eq-3} with variables from the algorithm,
whilst at the same time maintaining the if and only if satisfiability. The rest of the proof does
similar transformations, trying to keep the satisfiability of each system and also transforming
the expressions into something that looks like~\autoref{qcp-thm}.
